---
title: "EAR Pilot - Overview of FPMIS data for ESS"
author: "Michael Rahija, Josh Browning, EAR Team"
date: "October 15, 2015"
output: pdf_document
---

\section{Background}
This document shows a summary report of information contained in FPMIS for use in the ESS Activities Repository (EAR), AGSTATS, and a proposed Functional Technical Network for Agricultural Statistics (FTN). Through consultation with ESS colleagues and managers, the following information for EAR is required:

1. Project Title
2. Country (receipt)
3. In country partner institution
4. Status
5. Month year start
6. Month year finish
7. Country contact
8. FAO contact
9. Funding agency
10. Budget
11. Potential Synergies. 

It was suggested that some of this information may be readily available in FPMIS so some exploratory analysis of FPMIS data was undertaken. In fact all points except #11 and #3 can be extracted directly, however the data is of varying quality and indeed some follow-up with team capacity building focal points will be required. This report provides an overview of some results and tabulations of the FPMIS data, and draws conclusions on what can be used for EAR. 

A step-by-step approach was taken so that colleagues who are more familiar with FPMIS can make suggestions and comments on the tabulation and analysis of the information to improve its usefulness. 

\section{Data export, operationally active projects, geographic coverage}
First, a search was performed by Lead Technical Unit, **ESS**. **All** was selected as Project Status. .Below is a cross-tabulation showing number of projects by **Classification** and **Status**. 


```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 4, fig.width = 8}
#####################
##    CONFIGURE    ##
#####################

library(dplyr)
library(xlsx)
library(tidyr)
library(ggplot2)
library(RColorBrewer)
library(data.table)
library(maps)
library(countrycode)
library(scales)

sys <- Sys.info()

if(sys[7] == "josh"){
  dir = "~/Documents/Github/EAR_pilot/"
} else if(sys[5] == "x86_64"){
  dir = "~/Dropbox/FAO_ESS_STUFF/EAR_pilot" #Mac
} else if (sys[5] == "Michael"){
  dir = "C:/Users/Michael/Dropbox/FAO_ESS_STUFF/EAR_pilot"#HOME PC
} else if (sys[6]=="Rahija") {
  dir = "C:/Users/rahija/Dropbox/FAO_ESS_STUFF/EAR_pilot" #FAO PC
} else {
  stop("Implement location for current user!")
}

setwd(dir)


data <- read.csv("relevant_variables.csv", stringsAsFactors = FALSE)



## Clean data set, and filter for operationnaly active
source("R/cleanAndFilterFPMIS.R")
data.o <- cleanAndFilterFPMIS(data, ActiveOnly = FALSE)

##clean up project titles
data.o$ProjectTitle <- gsub("Ž","é", data.o$ProjectTitle)
data.o$ProjectTitle <- gsub("ci—n","ción", data.o$ProjectTitle)
data.o$ProjectTitle <- gsub("mise en .*uvre ","mise en ouvre ", data.o$ProjectTitle)


## Add staff column
data.o$staff <- paste(data.o$ProjectStaffBudgetHolder,
                      data.o$ProjectStaffLTOOfficer,
                      data.o$ProjectStaffLTUOfficer,
                      data.o$ProjectStaffProgrammeAssistance)


# data.o <- addTeam(data.o)

## Show overview of projects by status
tab <- data.o %>%
        group_by(ProjectStatus, ProjectClassification) %>%
        summarize(Number = length(ProjectTitle))
tab <- spread(tab, ProjectClassification, Number, fill = 0)

knitr::kable(tab, caption = "Number of Projects by Project Status and classification")
  
```

EAR, AGSTATS, and the FTN will focus exclusively on **Operationally Active** and capacity building projects in developing countries. Accordingly the data was filtered for **Operationally active** projects, that are classified as **Assistance to Member Countries**. There are only *78* of these projects in FPMIS. 

Team H (Census) has the most of projects by far with 32, Team G (Global Strategy and AMIS) has the second most with 16, and Team D (Food Security Statistics) has 12 projects. The regional offices (RAF, RAP, RLC, RNE, SAP, and SLC) have a combined total of 13 projects. *Note: Team was determined by project title, and staff. In 2 cases, the team coud not be determined.*


```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 4, fig.width = 8}
#filter data for only operationally active, and assistance
data.o <- filter(data.o, ProjectStatus == "Operationally Active" &
                   ProjectClassification == "Assistance to Member Countries")

source("R/addTeam.R")
data.o <- addTeam(data.o)

#data.o <- filter(data.o, !is.na(team))

temp <- data.o %>%
  group_by(team) %>%
  summarize(Projects = length(ProjectTitle))

temp.proj <- arrange(temp,desc(team))            

temp <- data.o %>%
  group_by(team, GeographicalCoverage) %>%
  summarize(Coverage = length(ProjectTitle))

temp.cov <- spread(temp, GeographicalCoverage,Coverage, fill = 0)


main <- merge(temp.cov, temp.proj, by = "team")
main <- arrange(main, desc(team))
colnames(main) <- gsub(" ","_",colnames(main))
main <- select(main, team, Global, Interregional, Regional, Sub_Regional,
               Country,not_specified, Projects)

ggplot(main, aes(x = team,y=Projects)) + 
  geom_bar(stat = "identity") +
  ggtitle("Operationally Active Projects in ESS") +
  xlab("Teams") +
  ylab("Number of Projects")

```

Then the number of projects were plotted by team and geographical coverage. By far, most of the projects are country level projects dominated by Team H, while only Teams K, J, I, G, and D have "Global" level projects. Team G projects account for the lion's share of Global and Regional projects. 

```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 4, fig.width = 8}

ggplot(temp, aes(x = GeographicalCoverage,y=Coverage, fill = team)) + 
  geom_bar(stat = "identity") +
  ggtitle("Operationally Active Projects in ESS by Team and Coverage") +
  xlab("Project Type") +
  ylab("Number of Projects") +
  labs(fill = "Team")
```

It may also be interesting to understand what kinds of projects certain teams have, as shown in the graph below.

```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 5, fig.width = 10}

ggplot(temp, aes(x = team, y=Coverage, fill = GeographicalCoverage)) + 
  geom_bar(stat = "identity", position = "fill") +
  ggtitle("Operationally Active Projects in ESS by Team and Coverage") +
  xlab("Team") +
  labs(fill = "Project Type") +
  scale_y_continuous("Percent of Projects", label = percent)
```

The focus of EAR will primarily be on country-level activities, and indeed for projects listed as *Country* in FPMIS, the country is provided in the system, but not for others. All of the countries that are beneficiaries of projects who coverage is marked *Country* are plotted on the map. There are 25 countries plotted with 1-3 projects in each.

```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 4, fig.width = 8}

mapData <- data.table(map_data(map = "world"))

## Repair polygons that wrap around the world
getNewGroup = function(d){
    n = nrow(d)
    newGroup = d[1:(n-1), group] == d[2:n, group] &
               abs(d[1:(n-1), long] - d[2:n, long]) > 180
    ## Don't switch right at the start
    newGroup = c(FALSE, newGroup)
    groupIndex = cumsum(newGroup)
    ## Each new group gets the old group number + 10,000 times the index
    d[, group] + 10000 * groupIndex
}
mapData$long[mapData$long >= 210] = mapData$long[mapData$long >= 210] - 360
invisible(mapData[, newGroup := getNewGroup(.SD), by = region])

summTab <- data.o %>%
    group_by(CountryName) %>%
    dplyr::summarize(Activities = n())

summTab$region <- summTab$CountryName
summTab$region = gsub(",( Federal Democratic | Democratic | the | The | United | )Republic (o|O)f", "",
                      summTab$region)
summTab$region = gsub(" , Kingdom of", "", summTab$region)

missingCountries <- merge(mapData, summTab, by = "region", all.y = TRUE)
missingCountries <- missingCountries %>% filter(is.na(lat))
## Check which countries are missing if running line by line:
# missingCountries

## Alternative approach using countrycode package: 
# temp$region = countrycode(temp$region, origin = "country.name",
#                           dest = "country.name", warn = TRUE)
# missingCountries <- merge(mapData, temp, by = "region", all.y = TRUE)
# missingCountries <- missingCountries %>% filter(is.na(lat))
# missingCountries

mapData <- merge(as.data.frame(mapData), summTab, by = "region", all.x = TRUE)
mapData <- mapData %>% arrange(order)

qplot(long, lat, fill = Activities, data = mapData, group = group,
     geom ="polygon", ylab="", xlab="") +
     scale_fill_continuous()
```

The table shows the project titles of Team G. The country beneficiary can in many cases be determinedby the Project Title; however, I suspect that this should be confirmed with the team focal point. See project for AMIS in the table. The project objectives specified in FPMIS can also be used. Anyway, the **determination of country beneficiaries can be automated for projects listed as "Country" for geographic coverage. The rest will require some follow-up with team focal points.** 

```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 4, fig.width = 8, results='asis', comment = FALSE}

data.nc <- filter(data.o, GeographicalCoverage != "Country" &
                          team == "Team G")

data.nc <- select(data.nc, GeographicalCoverage, ProjectTitle)



print(xtable::xtable(data.nc, align="lp{1.5in}p{4.5in}"),type = 'latex', include.rownames = FALSE, comment = FALSE)

```

\section{Donors}
For every project in FPMIS, a donor is listed. A table below shows the number of projects by donor. As demonstrated in the table, some are simply listed as "multilateral" with out a list of donors, and some others listed as multilateral show a single donor. 

```{r, echo=FALSE, message = FALSE, warning= FALSE, fig.height = 4, fig.width = 8}
temp <- data.o %>%
          group_by(Donor) %>%
          summarize(Projects = length(ProjectTitle))


knitr::kable(temp)
```

The **Original Budget** and **Donor** parameters in FPMIS can be used to compute the amount provided by each donor. In the case of multilateral donors if shares were entered into the system, they have been incorporated in the calculation. The Bill and Melinda Gates foundation has clearly commited the most with over 15 million followed by the European Union which has contributed almost 13 million.FAO is listed as the donor on project totalling 9.77 million. 

**This information seems to be missing some grants, possibly due to filtering. We'll follow-up!**

```{r, echo=FALSE, message = FALSE, warning= FALSE}
#create df only for budget analysis
data.d <- select(data.o, team, Donor, OriginalBudget)

#fix budget to make computations
data.d$OriginalBudget <- gsub("\\$","", data.d$OriginalBudget)
data.d$OriginalBudget <- gsub(",","", data.d$OriginalBudget)
data.d$OriginalBudget <- as.numeric(data.d$OriginalBudget)

#replace ambigious multilateral w/ "Multilateral"
multi.only.test <- grepl(" Multilateral \\(100%)",data.d$Donor,
                         ignore.case = TRUE)

data.d$Donor[multi.only.test] <- "Multilateral (not specified)"


######################################
#deal w/ non-ambigious multilateral  #
######################################
multi.test <- grepl("Multilateral: ",data.d$Donor)

#create master
master <- data.d[!multi.test,]
master$share <- rep(100,n=nrow(master))
master <- select(master, Donor, OriginalBudget,share)

#to extract and compute amounts
edit.list <- data.d[multi.test,]


edit.list$Donor <- gsub("Multilateral: ","",edit.list$Donor)

#split donors
temp.list <- strsplit(edit.list$Donor, ";")
temp.list <- unlist(temp.list)

#remove multi (100%)
test.list <- grepl("Multilateral \\(0%)",temp.list)
temp.list <- temp.list[!test.list]

#split number
temp.list <- strsplit(temp.list,"\\(")
temp.df <- data.frame(do.call("rbind",temp.list))
colnames(temp.df) <- c("donor", "share")

temp.df$share <- gsub("%)","", temp.df$share)

budgets <- c()
for(i in 1:length(temp.df$donor)){
  test.temp <- grepl(temp.df$donor[i],edit.list$Donor)
  
  budgets<- append(budgets,edit.list$OriginalBudget[test.temp])
  
}


master.multi <- cbind(temp.df,budgets)
colnames(master.multi) <- c("Donor","share","OriginalBudget")
master <- rbind(master.multi, master)

master <- mutate(master,total = (as.numeric(share)/100) * OriginalBudget)

master <- master %>%
          group_by(Donor) %>%
          summarize(total = sum(total)/1000000)


#clean up donor names
master$Donor <- as.character(master$Donor)

long0.test <- grepl("BMG",master$Donor)
master$Donor[long0.test] <- "BMGF"

long1.test <- grepl("combat",master$Donor)
master$Donor[long1.test] <- "UNCCD"

long2.test <- grepl("Commission for Africa", master$Donor)
master$Donor[long2.test] <- "UNECA"

long3.test <- grepl("World Food Programme", master$Donor)
master$Donor[long3.test] <- "WFP"

master$Donor <- gsub("^ ","",master$Donor)
master$Donor <- gsub(" $","",master$Donor)

ggplot(master,aes(x = reorder(Donor,total), y = as.numeric(total))) +  
  geom_bar(stat = "identity") +
  coord_flip() +
  xlab("Donors")+
  ylab("Sum of Original Budgets in million USD")
```

\section{Conclusions, and Next Steps}
The following are the primary conclusions:

1. FPMIS can provide most of the data required for EAR.
2. The country beneficiares of projects whose coverage is marked as "Country" are available in FPMIS. 
3. Some country beneficiaries of regional, interregional, etc. projects can be determined from the titles. For others, follow-up with team focal points will be required. 
4. FPMIS does not contain capacity building activities that are not standalone projects in FPMIS.
5. Donors information is not always available for multilateral projects. 
6. There are some descrepancies in FPMIS like projects for which the NTEs have passed, but are still classified as 'operationally active'. 
7. The donor and original budget needs to be reviewed for possible gaps.

The next steps:
1. Create EAR table for projects, and populate with information from FPMIS for 3 teams. See what follow-up is required.